---
title: "Average Temperature & CO2 Emissions in the US"
author: "Jacqueline Park, Jade D'Souza, Garam Park, Weicheng Wang"
date: "11/16/2019"
output: html_document
---

```{r setup, include=FALSE}
source("scripts/year_ave_temp_plot.R")
source("scripts/table.R")
source("scripts/finalmap.R")
source("scripts/pie_sources_emissions.R")
source("scripts/summary_info_temp.R")
library("knitr")
options(warn = -1)
```

Climate Change is something that isn't being given as much attention to in the United States. Our group's intention is to analyse the change in average temperature here, along with the change in CO2 emissions and try to find a correlation. 

We will be using two data sets in our project, one for the [average temperature](https://climateknowledgeportal.worldbank.org/download-data?fbclid=IwAR0_4L18xx27AdYQy2sccVFKefQomEvtQvtsSnvgdLyUJdQnMbtho91XDig) seen in the US per month of every year, along with [CO2 emissions](https://data.worldbank.org/indicator/en.atm.co2e.pc) seen per country. We plan on filterng our the CO2 emissions for just data seen in the US for most of our analysis.

> Our CO2_emissions file has information on the sources from which these CO2 emissions come from (ie residential buildings, manufacturing sites), so we will use our information_df.csv file to help us read the information in our CO2_emissions.csv. Our source for this information is the same source we used for CO2_emissions.csv, as hyperlinked in the above paragraph.


## Trends Seen in Temperature Depending on Year

```{r, echo = FALSE}
average_temp_plot
```
This line plot helps further visualize this, and is intended to further show increase in temperature seen in recent years. There's a gradual rise in temperature from 1991. There was a slight dip from the years 2002-2009, but the increase in temperature only got worse from thereon, but we can clearly see average temperature increasing from 2010. Looking at our outliers, 2016 still was the hottest, with 2015 being a close second. This helps in proving that there has been an increase in temperature in the US, and we hope to find a correlation between CO2 emissions and temperature with more analysis. 


## Conclusions on Yearly Temperatures
More specifically, the highest mean temperature seen from 1991 is `r list_avg_temp$max_mean`. We expected that this would be seen in `r list_avg_temp$max_year` as our data consistently shows us that this has been the hottest year. The lowest mean temperature was `r list_avg_temp$min_mean`, occuring in `r list_avg_temp$min_year`. From this data too, we can conclude that temperature has been rising over the last `r list_avg_temp$years` with a total rise in average of `r list_avg_temp$year_max_mean`. This helps answer the first part of our second question -- 2016 was the hottest year.


## CO2 Emissions and Temperature
```{r, echo = FALSE}
kable(temp_emission,
      row.names = FALSE,
      col.names = c("Year", "Total CO2 Emissions", "Change in Temperature"))
```


Our table above displays the yearly temperatures along with the CO2 emissions seen and the temperature in that year. This would help us in understanding and answering our question of if there is a correlation seen between emissions and temperature. Our table is sorted by total CO2 emissions, for ease in looking at the years with the worst CO2 emissions. Our data consists of the years 1998-2012. 2012, the most recent year, has the highest change in temperature with the least amount of total emissions. 2009, the year with the second least change in temperature is also the year with the 3rd least total emissions of CO2. This helps us in answering our first question -- if there is a correlation between rise in average temperature and CO2 emissions. The linear regression is `r regression`, so there doesn't seem like there's much of a correlation between the temperature change and total CO2 emissions seen per year, as the data from the table suggests. To answer the second question, our hottest year is not the year with the most CO2 emissions.


## Comparison of Personal Expenditures & CO2 emissions  
```{r, echo=FALSE}
personal_expenditure_2000
personal_expenditure_2015
```
Comparing the total consumption of fuel from personal expenditure in 2000 and 2015, we can see that most consumption has varied by about 2%, except for residential, commercial and public services. The greatest increase in fuel consumption in these 2 years is transport, which could vary over a multitude of reasons such as an increase in [air travel](https://www.nationalgeographic.com/environment/urban-expeditions/transportation/air-travel-fuel-emissions-environment/) and a decrease in use of [public transport](https://www.economist.com/leaders/2018/06/23/how-to-stop-the-decline-of-public-transport-in-rich-countries) in America. The intention of using a pie chart was to compare different years and easily look at categories that consume more fuel than others. In this case, electricity and heat consistently are the highest categories in fuel consumption, almost occupying half of the pie chart. This helps us in answering our third question, being where does most fuel consumption come from.


## CO2 Emissions (kt) around the World

```{r, echo=FALSE}
world_map
```
Our group was curious to see how each country differed from each other with their changes in CO2 emissions over 17 years. This chart was intended to see if other countries outperform others in terms of their CO2 emissions, or if all countries produced around the same amounts of CO2. As we can see extremely clearly, China has had the highest amount of change in CO2 emissions (6.96M), with India having the second highest(1.30M). We can also observe that Middle Eastern & Asian countries tend to have a greater change in these emissions compared to other countries. These changes in CO2 emissions (in India & China particularly) would probably also explain them being countries with [a notable amount of air pollution.](http://worldpopulationreview.com/countries/most-polluted-countries/). From this, we can also see that an outlier is the United States (because of how light its color is in comparison with other countries) with -146.73k of change.
